Deep Reinforcement Learning for Conservation Decisions

### 2021-06-08 manuscript/manuscript.Rmd

Neural networks become useful in RL when the environment has a large state-action space, which happens frequently with realistic decision-making problems. 
Classic RL algorithms, like policy iteration, value iteration and TD-learning, fail when the state-action space is large, e.g. when the possible states and actions can not be represented in a tractable table [@suttonbarto].
Yet, over the last 10 years, researchers have shown that neural network-based RL algorithms can perform well on previously unsolvable environments [@DQN, @alphaGo2016].
Neural networks have been widely useful in reinforcement learning because neural nets have the property of being general function approximators [@universalapproximators].
Whenever there is a need for an agent to approximate some function, say a policy function or value function, neural networks can be used in this capacity.
While there are other function approximators that can be used in RL, e.g. Gaussian processes [@grande14], neural networks have excelled in this role because of their ability to learn complex, non-linear, high dimensional functions and their ability to adapt given new information [@briefsurveydrl].
We will not discuss in detail how neural networks work [see @lecun_deep_2015], but the general process for how neural networks are used in RL is that at each learning step, the agent, will adjust the parameters in its neural network with which it approximates the policy function and/or value function, so that according to past experiences, the agent will be more likely to engage in highly rewarding behavior in the future.

Over the course of training, RL agents will engage in trial and error learning, whereby it will be necessary for the agent to explore new sequences of actions as well as exploit past high rewarding sequences.
Occasionally, the agent will need to propose exploratory actions that may result in higher rewards than previously seen. 
The agent may also need to focus on past sequences of actions which may require marginal improvement to reach optimality.
It is very difficult for the agent to know when it should engage in explorative or exploitative behavior. 
This *exploration-exploitation dilemma* is an open research problem [@berger-tal_exploration-exploitation_2014]. 
Current deep RL algorithms take a range of approaches towards this (see  Appendix A).
The manner in which RL algorithms explore depend on whether the algorithms are *on-policy* or *off-policy*.
The distinction between on and off-policy evaluation comes from what policy the agent is using to select actions during training.
For off-policy algorithms, the agent uses a policy during training that can be different from the policy that the agent will use during evaluation; while in on-policy algorithms, the agent updates the same policy in training as the agent will use during evaluation.
Other points of divergence include whether the agent learns a deterministic or stochastic policy, and
whether the algorithm is formulated to work on a discrete, continuous or both a discrete and continuous state and action space.

@@

For illustrative purposes, we consider the simplest version of the dynamic optimal harvest problem as outlined by [@Clark1973] (for the deterministic case) and [@Reed1979] (under stochastic recruitment).
The manager seeks to optimize the net present value (discounted cumulative catch) of a fishery, observing the stock size each year and setting an appropriate harvest quota in response.
In the classical approach, the best model of the fish population dynamics must first be estimated from data, potentially with posterior distributions over parameter estimates reflecting any uncertainty.
From this model, the optimal harvest policy -- that is, the function which returns the optimal quota for each possible observed stock size -- can be determined either by analytic [@Reed1979] or numerical [@Marescot2013] methods, depending on the complexity of the model.
In constrast, a model-free deep RL algorithm makes no assumption about the precise functional form or parameter values underlying the dynamics -- it is in principle agnostic to the details of the simulation.

We illustrate the deep RL approach using the algorithm known as Twin Delayed Deep Deterministic Policy Gradient or more simply, TD3 [@TD3].
TD3 is an extension of Deep Deterministic Policy Gradient [DDPG, @DDPG], which itself extended the original Deep Q-Network [DQN, @DQN] algorithm to continuous action space.
A step-by-step walk-through for training agents on this environment is provided in the Appendix.
We compare the resulting management, policy, and reward under either RL agent to that achieved by the optimal management solution [Fig 2].
Despite having no knowledge of the underlying model, the RL agent learns enough to achieve nearly optimal performance.

@@

The cumulative reward (utility) realized across 100 stochastic replicates is indistinguishable from that of the optimal policy [Fig 2].
Nevertheless, comparing the mean state over replicate simulations reveals some differences in the RL strategy, wherein the stock is maintained at a slightly higher-than-optimal biomass.
Because our state space and action space are sufficiently low-dimensional in this example, we are also able to visualize the policy function directly, and compare to the optimal policy [Fig 2].
This confirms that quotas tend to be slightly lower than optimal, most notably at larger stock sizes.
These features highlight a common challenge in the design and training of RL algorithms.
RL cares only about improving the realized cumulative reward, and may sometimes achieve this in unexpected ways.
Because these simulations rarely reach stock sizes at or above carrying capacity, these larger stock sizes show a greater deviation from the optimal policy than we observe at more frequently visited lower stock sizes.
Training these agents in a variety of alternative contexts can improve there ability to generalize to other scenarios.

How would an RL agent be applied to empirical data?
In principle, this is straight forward.
Once our agent has been trained on appropriate simulation environments, it is able to propose a quota given an observation of the stock after appropriate transformation of variables.
To illustrate this, we examine the quota that would be recommended by our newly trained RL agent, above, against historical harvest levels of Argentine hake based on stock assessments from 1986 - 2014 [@ram, see Appendix D].
Hake stocks showed a marked decline throughout this period, while harvests decreased only in proportion [Fig 3].
In contrast, our RL agent would have recommended significantly lower quotas over most of the same interval, including the closure of the fishery as stocks were sufficiently depleted.
While we have no way of knowing for sure if the RL quotas would have led to recovery, let alone an optimal harvest rate, the contrast between those recommended quotas and the historical catch is notable.

@@

This approach is not as different from conventional strategies at it may seem.
In a conventional approach, ecological models are first estimated from empirical data, (stock assessments in the fisheries case).
Quotas can then be set based directly on these model estimates, or by comparing alternative candidate "harvest control rules" (policies) against model-based simulations of stock dynamics.
This latter approach, known in fisheries as Management Strategy Evaluation [MSE; @Punt2016] is already very closely analogous to the RL process.
Instead of researchers evaluating a handful of control rules, the RL agent proposes and evaluates a plethora of possible control rules autonomously.

>>>>

Neural networks become useful in RL when the environment has a large state-action space, which happens frequently with realistic decision-making problems. 
Classic RL algorithms, like policy iteration, value iteration and TD-learning, fail when the state-action space is large, e.g. when the possible states and actions can not be represented in a tractable table [@suttonbarto].
Yet, over the last 10 years, researchers have shown that neural network-based RL algorithms can perform well on previously unsolvable environments [@DQN, @alphaGo2016].
Neural networks have been widely useful in reinforcement learning because neural nets have the property of being general function approximators [@universalapproximators].
Whenever there is a need for an agent to approximate some function, say a policy function or value function, neural networks can be used in this capacity.
While there are other function approximators that can be used in RL, e.g. Gaussian processes [@grande14], neural networks have excelled in this role because of their ability to learn complex, non-linear, high dimensional functions and their ability to adapt given new information [@briefsurveydrl].
We will not discuss in detail how neural networks work -- see [@lecun_deep_2015] for further background --, but the general process for how neural networks are used in RL is that at each learning step, the agent, say a policy gradient-based agent, will adjust the parameters in its policy-approximating neural network so that according to past experiences, the agent will be more likely to engage in highly rewarding behavior in the future. 
This same process occurs in value learning and actor-critic-based deep RL algorithms, except these agents are adjusting either a value network or both a value and policy network respectively

On top of the aforementioned distinctions of deep RL algorithms, there are more algorithms design choices that are made. 
A very significant consideration is deciding how the agent will explore the environment.
Another is whether the algorithm can work on a discrete, continuous or both a discrete and continuous state and action space.
In Appendix A, we provide more detail on the important components of RL algorithms.

@@

For illustrative purposes, we consider the simplest version of the dynamic optimal harvest problem as outlined by [@Clark1973] (for the deterministic case) and [@Reed1979] (under stochastic recruitment).
In short, the manager seeks to optimize the net present value (discounted cumulative catch) of a fishery, observing the stock size each year and setting an appropriate harvest quota in response.
In the classical approach, the best model of the fish population dynamics must first be estimated from data, potentially with posterior distributions over parameter estimates reflecting any uncertainty.
From this model, the optimal harvest policy -- that is, the function which returns the optimal quota for each possible observed stock size -- can be determined either by analytic [@Reed1979] or numerical [@Marescot2013] methods, depending on the complexity of the model.
However, unlike an analytic or SDP solution that is derived directly from the estimated model, a model-free deep RL algorithm makes no assumption about the precise functional form or parameter values underlying the dynamics -- it is in principle agnostic to the details of the simulation.

Here, we apply a version of the deep RL algorithm known as Twin Delayed Deep Deterministic Policy Gradient or more simply, TD3 [@TD3].
A step-by-step walk-through for training both agents on this environment is provided in the Appendix.
We compare the resulting management, policy, and reward under either RL agent to that achieved by the optimal management solution [Fig 2].
Despite having no knowledge of the underlying model and clear differences in the resulting strategies, both RL agents learn enough to achieve nearly optimal performance.

@@

These features highlight a common challenge in the design and training of RL algorithms.
RL cares only about improving the realized cumulative reward, and may sometimes achieve this in unexpected ways.
Because these simulations rarely reach stock sizes at or above carrying capacity, these larger stock sizes show a greater deviation from the optimal policy than we observe at lower stock sizes.
As intuition would suggest, it is possible to devise alternative training schemes which may do a better job sampling the simulation dynamics at high stock levels.
Training these agents in a variety of alternative contexts can improve there ability to generalize to other scenarios.

<!-- 
Add text about empirical example
-->

@@
